\LARGE \bf Human-Robot Interaction: A Deliberative Architecture

### 2013-07-25 paper.tex
Minor tweak
Work on the related work/discussion section
Added many new references, reorganized the section

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

@@

\subsection{Building the Deliberative Layer}

@@

\section{Perspectives}
\label{sect|conclusion}

\subsection{Contributions and Related work}\label{sec:soa}

The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhav2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasize that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realize it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Tambe2005a,Breazeal2003} and they
all have shown benefits of this approach. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way. Finally, today only few
systems~\cite{Fong_2006,Breazeal2003,Sisbot2008} take humans into
account at all levels.

Perspective Taking is a human ability which allows one to put
him/herself in another person's point of view. Studied in
psychology literature~\cite{Flavell1992,Tversky1999}, this ability is
crucial when interacting with people by allowing one to reason on
others' understanding of the world in terms of visual perception, spatial
descriptions, affordances and beliefs, etc.
Therefore, in the last years these notions have been gradually
employed in Human-Robot Interaction.~\cite{Breazeal2006} presents a
learning algorithm that takes into account information about a
teacher's visual perspective in order to learn a
task. ~\cite{Johnson2005} apply visual perspective taking for action
recognition between two robots.~\cite{Trafton2005} use both visual and
spatial perspective taking for finding out the referent indicated by a
human partner.

Spatial reasoning~\cite{O'Keefe1999}, on the other hand, has been used
for natural language processing for applications such as direction
recognition ~\cite{Kollar2010,Matuszek2010} or language
grounding~\cite{Tellex2010}.~\cite{Skubic2004} presented a spatial
reasoner integrated in a robot which computes symbolic positions of
objects.

@@

Those items are not new \emph{per se}. It seems however interesting to
underline the shift of focus this implies during the design and integration
phases.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{abstract}

@@

\subsection{Building a human-aware deliberative layer}

@@

\section{Discussion: when AI enables HRI}
\label{sect|conclusion}

This last section goes over again the main contributions we have presented
so far: why are they relevant to robotics and artificial intelligence? How do
they compare to similar approaches? How do we justify our technical choices and
what remain to be developed/explored in the future?

\subsection{Foundations of cognition for the robot}

Because we build autonomous \emph{companion} robots, designed to live and
interact in environments primarily conceived for humans, we endow our machines
with a set of fundamental cognitive capabilities. The robots perform typically
in dynamics, partially unknown environments, with a variety of places,
artifacts and agents that may carry complex semantics.

While our main focus is on the cognitive abilities required to interact with
humans, \emph{situation assessment}, \emph{spatial reasoning}, \emph{knowledge
representation and reasoning}, \emph{task planning} or \emph{high-level
control} are mandatory cognitive foundations that we first have to mention.

\paragraph{Situation assessment, spatial reasoning}

Real service robots need to ground their decision-making processes in their
perceptions, be it physical (\eg what a camera sees) or not (\eg remote
knowledge bases, dialogue). We call \emph{physical situation assessment} the
cognitive skill a robot exhibits when it assesses the nature and content of its
surroundings and efficiently monitors its evolution.

Numerous approaches exist, like amodal \emph{proxies}~\cite{Jacobsson2008},
grounded amodal representations~\cite{Mavridis2006}, semantic maps
(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planing and object classification~\cite{Lorken2008,
Varadarajan2011}.

Our approach, implemented in the {\sc Spark} module, is a grounded amodal
representation of the robot environment, where different sensing modalities
(vision-based fiducial markers tracking, RGBD human tracking, motion capture)
are merged and where amodal spatial reasoning takes place (explanation
generation for unexpected perceptions like a disappearing object).

Spatial reasoning~\cite{O'Keefe1999} is a field in its own right, and has been
used for natural language processing for applications such as direction
recognition ~\cite{Kollar2010,Matuszek2010} or language
grounding~\cite{Tellex2010}.~\cite{Skubic2004} presented a spatial reasoner
integrated in a robot which computes symbolic positions of objects.

One of the main limit of our current approach is the way semantics are
conveyed: because we use 2D markers for artifacts, we skip most of the object
extraction and recognition challenges. Each object has a unique identifier,
which enable us to load a proper CAD model suitable for 3D spatial reasoning
and prevent recognition mistakes in the knowledge base. While {\sc Spark}
algorithms themselves are ignorant of the input sources, and would work equally
well with a full object recognition stack, it is not currently supported.

\fxwarning{spatial reasoning -> mention we do not support extensive temporal reasoning}

\paragraph{Robot control}

\fxwarning{Robot control related work -> what is the originality of SHARY? why is it relevant from an AI point of view? (Mathieu, Aurelie)}

Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture called
\textsc{CRAM} (Cognitive Robot Abstract Machine) that integrates
\textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
Prolog. Its underlying storage is based on an OWL ontology, derived from
\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KnowRob} have been demonstrated on
several real-world scenarios, where natural language recipes extracted from the
Internet had to be translated into plans and executed in a kitchen environment,
perceived and rebuilt on-line by the robots. While Prolog offers more flexible
modeling (no constraints on the arity of predicate, where Description Logics as
used in our work are limited to binary predicates), it is based on the closed
world assumption (if something cannot be inferred to be true, it is inferred to
be false) whereas we rely on the open world assumption, which is more realistic
in real world scenarios. A probabilistic extension of \textsc{KnowRob}, called
\textsc{ProbCog}~\cite{Jain2009} is also available. While in principle
possible, currently the CRAM architecture does not provide explicit support for
interacting with humans.

\paragraph{Knowledge model}

\subsection{Interacting with humans}

\paragraph{Building a model of a human during the interaction} Perspective
Taking is a human ability which allows one to put him/herself in another
person's point of view. Studied in psychology
literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
interacting with people by allowing one to reason on others' understanding of
the world in terms of visual perception, spatial descriptions, affordances and
beliefs, etc.  Therefore, in the last years these notions have been gradually
employed in Human-Robot Interaction.~\cite{Breazeal2006} presents a learning
algorithm that takes into account information about a teacher's visual
perspective in order to learn a task. ~\cite{Johnson2005} apply visual
perspective taking for action recognition between two
robots.~\cite{Trafton2005} use both visual and spatial perspective taking for
finding out the referent indicated by a human partner.

\fxwarning{What are our contribution regarding perspective taking? (Mathieu)}
\fxwarning{Mention other source of knowledge about humans -- including social rules for instance}

\paragraph{Reasoning with the humans}

\fxwarning{Mention we represent the human knowledge as well (Severin)}

\paragraph{Multi-modal communication}

\subsection{Joint actions with humans}
\label{sec:soa}

The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasize that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realize it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Tambe2005a,Breazeal2003} and they
all have shown benefits of this approach\fxwarning{which approach? how it is related to larger AI challenges?}. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way\fxwarning{is it smthg we improve one?}. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels\fxwarning{Old references, no new ones? we should also say what is new/better with our approach (Rachid)}.

@@

Those items are not new \emph{per se}. It seems however interesting to
underline the shift of focus this implies during the design and integration
phases: components of our deliberative layer are defined and glued together by
the knowledge they produce and consume.
